The LSST-DESC 3x2pt Tomography Optimization Challenge

### 2021-03-24 paper/paper.tex
edited abstract and left comments through sec 5.2

\begin{document}
\title{The LSST DESC 3x2pt Tomography Optimization Challenge}
\input{authors}

\begin{abstract}
In this paper we present the results of the DESC 3x2pt tomography challenge,
which aimed to compare strategies for optimizing the photometric binning required
for the main LSST 3x2pt analysis. This task is made particularly delicate in the
context of a metacalibrated lensing catalog, as only the photometry from the
bands  included in the metacalibration process (riz and potentially g) can be
used to define this tomographic binning. Using a set of realistic simulations
generating galaxy redshifts and simulated photometry, the goal of the
challenge was to propose bin assignment strategies that maximized various
figures of merit for a 3x2pt analysis. This challenge was highly idealized: it ignored
spectroscopic completeness and limited size for the training sample, but was complex enough
to address the main question of binning optimization. Based on the challenge
dataset, we review and compare the different methods that have been proposed by
participants. We find that even from this limited photometry information,
various methods are able to yield binning schemes that achieve high figure-of-merit scores.
We further find that the g band adds around 15\% to typical figures of merit, and
that there is significant value in optimizing different bin assignments for specific
science cases.
\end{abstract}

@@

@@

Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
binning galaxies by redshift.
This approach captures almost all the available information in lensing data, since it measures
an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
gains from moving beyond tomographic methods.}.  Binning galaxies by approximate redshift also lets us 
model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
While fully 3D methods have been proposed, prototyped, and shown to have significant promise, 
\citep{heavens,kitching}, the tomographic approach remains the standard within the field.

We are free to assign galaxies to different tomographic bins in any way we wish; changing the choice
can potentially the ease with which we can calibrate
the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with peaks at different redshifts if that is useful.
We should choose, then, an assignment algorithm that maximises the constraining power for a science
case of interest.  This challenge explores such algorithms.

Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
\citep{euclid,roman}. The general question of optimizing tomography was recently discussed in detail
in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges.

In this paper we discuss and evaluate the related challenge
of using a limited set of colour bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  
A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

The Dark Energy Science Collaboration (DESC) Tomographic Challenge was launched in May 2019 and 
advertised among the LSST cosmology community, though it was open to outside entrants.

In the challenge, participants assigned simulated galaxies to tomographic bins,
using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
information in the final split data set; this is generally acheived by separating different bins by 
redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
redshift bin edges from their training sample, but in either case had to classify galaxies into the 
different bins.

@@

The challenge was open to any entrants, not just those already involved in DESC, but it was
advertised primarily through Rubin, and so is perhaps best described as semi-public.
Participants tested and ran their methods locally, but submitted code to be run as part of a central
combined analysis on the NERSC supercomputing facility.  No prize was offered apart from recognition.

@@

We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
validation, and testing components.  In each case participants were supplied with the training and 
validation data, and the testing data was kept secret for final evaluation. In each case the three
data sets were random sub-selections of the full data sets, with the training and testing sets 25\% of 
the full data each and the validation 50\%. In practice many entrants trained on a subset of
this, so to make the final evaluation phase tractable we trained on 1M objects for both data sets.

We used the second Buzzard catalog for two reasons.  First, to check the significance of
a limitation of galaxy colour assignments in CosmoDC2, described below.  And second, to check
for over-fitting in the design of models themselves, and thus to determine whether two
reasonable but different galaxy populations can be optimized by the same methods; this will
tell us whether our conclusions might be reasonable for real data. We discuss correlations
between metric scores on the two data sets in Section \ref{sec:metric-results}.

Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
and sizes.  We simulate and add noise to the objects as described below.

@@

The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.  It used dark
matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
an algorithm that matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
using the LSST band passes were provided in the catalog.

@@

\subsubsection{Post-processing} \label{sec:data-proc}

@@

The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.

\begin{figure}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
after the initial cuts described in the text.}
\label{fig:initial_nz}
\end{figure}

\begin{figure}[htbp]
% \centering
\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
available to participants.}
\label{fig:colour_colour}
\end{figure}

\subsection{Metrics}
\label{sec:metrics}

@@

Metric 1 would be approximately measurable on real data without performing a full cosmology analysis,
whereas metric 2 would be the final output of such an analysis.  They therefore test performance at what
would be multiple stages in the wider pipeline.

\subsubsection{Metric 1: Sigal-to-noise}
Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a Gaussian estimate of the covariance between them, as described in
\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

\subsubsection{Metric 2: Figure-of-merit}

@@

The challenge was structured as a python library, in which each classification
method was a class.  The abstract base parent class, {\sc Tomographer}, 
and other challenge machinery, performed initialization,
configuration option handling, data loading, and computing derived quantities
such as colours from magnitudes.

@@

Two control methods were used to test the challenge machinery and metrics.
The first method randomly assigned galaxies to each tomographic bin, resulting
in bins which each had the same $n(z)$, on average (the random noise fluctuations
are very small when using this number of galaxies).  Using this method, we expect
that increasing the number of bins should not increase any metric score, since bins are
maximally correlated.  We find this to be true for all our metrics.

@@

\section{Submitted Methods} \label{sec:methods}

Twenty five methods were submitted to the challenge, including the control methods described above.
Most used one machine learning method of one type or another to perform the primary classification,
rather than trying to perform a full photometric redshift analysis.

These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
% \setlength{\itemsep}{0pt}%
\begin{description*}
	\item[\textsc{Random}] the random assignment control method (organizers)
	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
	\item[\textsc{UTOPIA}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
	\item[\textsc{FFNN}] a feed-forward neural network (EN)
	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
\end{description*}

\section{Results \& Discussion} \label{sec:results}

@@

The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).

\subsection{Overview}

The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.

In all these tables, missing entries are shown for two reasons.
The first, marked with an asterisk, is when a method did not run for the given scenario,
either due to memory overflow or taking an extremely long time.
The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
small number count, such that the spectrum could not be calculated.

The challenge did not require that entrants clasified all galaxies; ambiguous
galaxies could be discarded at the cost of number density.  Several methods
took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges.

\begin{table*}[]
\begin{tabular}{|l|llll|llll|}
\hline
                & \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
\hline
\end{tabular}
\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
CosmoDC2 part of the challenge.}
\label{tab:cosmodc2}
\end{table*}

The overall winning method in the 3x2pt case in the fiducial griz 3x2pt CosmoDC2 metric was \textsc{funbins},
which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin ugrizy bands.
It is clear from this that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded, at least in our idealised scenario.
In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

There is a clear widening of the $n(z)$ when losing the g-band.  This widening leads to a loss in 
constraining power, which is discussed more in Section \ref{sec:gband}.

For comparison, we also show results using all six LSST bands, ugrizy; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
is 169.7, barely higher than the score for griz, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.

@@

\begin{figure}[htbp]
% \centering
\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
generating nine bins with the CosmoDC2 data.
The upper panel shows results for the method applied to riz bands, and the middle on griz. The
increased spread and overlap of the bins in the middle panel decreases the overall constraining
power and signal-to-noise.  The lower panel is for comparison and is not part of the challenge,
and illutrates results using all the ugrizy bands.}
\label{fig:funbin_nz}
\end{figure}

@@

\section{Mistakes} \label{sec:mistakes}
The challenge organizers made a number of mistakes when building and running the challenge.
These do not invalidate our process or results, but we describe some of them here in the hope of
being useful for future challenge designers.

>>>>

\begin{document}
\title{The LSST-DESC 3x2pt Tomography Optimization Challenge}
\input{authors}

\begin{abstract}
This paper presents the results of the LSST-DESC 3x2pt tomography challenge, which served as a first step toward optimizing the tomographic binning strategy for the main LSST 3x2pt cosmology analysis. 
The task of choosing an optimal tomographic binning scheme for a photometric survey is made particularly delicate in the context of a metacalibrated lensing catalog, as only the photometry from the bands included in the metacalibration process (traditionally, {\em riz} and potentially {\em g}) can be used in sample definition.
The goal of the challenge was to collect and compare bin assignment strategies under various metrics of a standard 3x2pt cosmology analysis in a highly idealized setting to establish a baseline for realistically complex follow-up studies; 
in this preliminary study, we used two sets of cosmological simulations of galaxy redshifts and photometry under a simple noise model neglecting outliers and variation in observing conditions, and contributed algorithms were provided with a representative and complete training set.
We review and evaluate the entries to the challenge, finding that even from this limited photometry information, multiple algorithms yield binning schemes with high figure-of-merit scores.
We further find that the {\em g} band improves metric performance by $\sim15\%$ and confirm that the optimal bin assignment strategy depends on the specific science case.
\end{abstract}

@@

% AIM: Suggestion: Use a macro for DESC because it shows up with and without expansion all over

@@

% AIM: The next two paragraphs should present tomographic binning as an experimental design choice up front.

Most lensing and 3x2pt analyses have been performed \emph{tomographically}, 
binning galaxies by redshift.
% AIM: I get what this paragraph is saying, but it's worded ambiguously and can read as almost its exact opposite. I suggest rewording to specify more of the instances of "this," "that," and "it."
This approach captures almost all the available information in lensing data, since it measures
an integrated effect and so galaxies nearby in redshift probe very similar fields.  For photometric
foreground samples it is similarly near-lossless, since redshift estimates of such galaxies have
large uncertainties\footnote{Spectroscopic foreground samples may be more likely to see significant 
gains from moving beyond tomographic methods.}.  Binning galaxies by approximate redshift also lets us 
model galaxy bias, intrinsic alignments, and other systematic errors en masse in a more tractable way.
While fully 3D methods have been proposed, prototyped, and shown to have significant promise, 
\citep{heavens,kitching}, the tomographic approach remains the standard within the field.

% AIM: typo in the next sentence {potentially the ease}
We are free to assign galaxies to different tomographic bins in any way we wish; changing the choice
can potentially the ease with which we can calibrate
the bins, but any choice can lead to correct results. The bins need not even be purely tomographic: we can happily correlate bins with peaks at different redshifts if that is useful.
We should choose, then, an assignment algorithm that maximises the constraining power for a science
case of interest.  This challenge explores such algorithms.

% AIM: Move the following sentence two paragraphs up to where tomography is introduced
Tomographic 3x2pt measurements will be a key science goal in the upcoming \emph{Stage IV} surveys,
including the Rubin Observatory \citep{rubin} and the Euclid and Roman space telescopes
\citep{euclid,roman}. The general question of optimizing tomography was recently discussed in detail
in \citet{rainbow} using a self-organizing map approach to target up to five tomographic bins.  For this
configuration they find that equally spaced redshift bins are a good proxy
for an optimized method, but note, importantly, that we should not be constrained to trying to directly
match our bin selections to specific tomographic bin edges.
% AIM: Above sentence should cite other strategies as well.

% AIM: insert comma {paper we}
% AIM: messaging -- establish scope with the basic info now in the abstract
In this paper we discuss and evaluate the related challenge
of using a limited set of colour bands to make those assignments, motivated by requirements of using the
metacalibration method for galaxy shape measurement to limit biases.  
A similar question, with different motivations, was explored in \citet{jain}, who find that 3-band 
gri tomography could be effective for tomographic selection, a result we will echo here in a new context: in this paper we describe the results of a 
challenge using simulated Rubin-like data with a range of methods submitted by different teams.  We explore how many tomographic bins we can effectively generate using only a subset of bins, and compare methods submitted to the challenge.

@@

% AIM: 2020, not 2019!
The Dark Energy Science Collaboration (DESC) Tomographic Challenge was launched in May 2019 and 
advertised among the LSST cosmology community, though it was open to outside entrants.

In the challenge, participants assigned simulated galaxies to tomographic bins,
using only their (g)riz bands and associated errors, and the galay size.  Their goal was to maximize the cosmological
information in the final split data set; this is generally acheived by separating different bins by 
redshift as much as possible.  Particpants were free to optimize or simply select nominal (target) 
redshift bin edges from their training sample, but in either case had to classify galaxies into the 
different bins.
% AIM: We must explicitly state the assumptions that all methods were expected to be ML/training-based and that this is framed as a classification problem, and explain how that fits with the paradigm of how people think about tomography, which is traditionally framed as a very different redshift regression problem.

@@

% AIM: Was it actually advertised in Rubin beyond DESC at all?
The challenge was open to any entrants, not just those already involved in DESC, but it was
advertised primarily through Rubin, and so is perhaps best described as semi-public.
% AIM: "tested and ran" --> "developed and tested" 
% AIM: remove comma {locally, but}
% AIM: Final tests used to generate results on secret test set were conducted on NERSC, no?
Participants tested and ran their methods locally, but submitted code to be run as part of a central
combined analysis on the NERSC supercomputing facility.  No prize was offered apart from recognition.
% AIM: Don't specify NERSC up here before it's defined in Sec. 3.4.

@@

\begin{figure}[htbp]
	% \centering
	\includegraphics[width=0.9\linewidth]{results/initial_data.pdf}
	% AIM: use alpha \neq 1
	\caption{The underlying number density $n(z)$ for the two catalogs used in the challenge,
		after the initial cuts described in the text.}
	% AIM: specify /redshift/ number density
	\label{fig:initial_nz}
\end{figure}

\begin{figure}[htbp]
	% \centering
	\includegraphics[width=0.9\linewidth]{results/colour_colour.pdf}
	% AIM: use alpha \neq 1
	\caption{A colour-colour diagram for the two catalogs used in the challenge, showing the bands
		available to participants.}
	\label{fig:colour_colour}
\end{figure}

% AIM: Rest of paper uses past tense so "run" --> "ran"
% AIM: Train/Validate/Test set splitting should be described in post-processing rather than preamble, and assumption of ML/training-based approach should be stated explicitly in preamble to entire section.
We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
validation, and testing components.  In each case participants were supplied with the training and 
validation data, and the testing data was kept secret for final evaluation. In each case the three
data sets were random sub-selections of the full data sets, with the training and testing sets 25\% of 
the full data each and the validation 50\%. In practice many entrants trained on a subset of
this, so to make the final evaluation phase tractable we trained on 1M objects for both data sets.

% AIM: We don't have to justify the second dataset unless we provide a reason why the first was insufficient. . . the presentation here needs some reworking.
We used the second Buzzard catalog for two reasons.  First, to check the significance of
a limitation of galaxy colour assignments in CosmoDC2, described below.  And second, to check
for over-fitting in the design of models themselves, and thus to determine whether two
reasonable but different galaxy populations can be optimized by the same methods; this will
tell us whether our conclusions might be reasonable for real data. We discuss correlations
between metric scores on the two data sets in Section \ref{sec:metric-results}.

% AIM: typo {Each of these data set}
Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
and sizes.  We simulate and add noise to the objects as described below.

@@

% AIM: Say it was also used for DESC DC1 studies, cite e.g. PZ DC1 paper
The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.  It used dark
matter simulations from L-GADGET2 \cite{gadget2} and then added galaxies using
an algorithm that matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3 (significantly shallower than CosmoDC2)
and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
using the LSST band passes were provided in the catalog.

@@

\subsubsection{Post-processing} \label{sec:data-proc}
% AIM: Put the info from CosmoDC2/Buzzard pertaining to status after post-processing steps here instead.

@@

% AIM: specify /redshift/ number density and provide n(z) symbol because it is used without introduction later
The overall number density of the two data sets is shown in Figure \ref{fig:initial_nz},
and a set of colour-colour diagrams  in Figure \ref{fig:colour_colour}.
% AIM: If submitting to AAS journal, colour --> color; if submitting to MNRAS, catalog --> catalogue

\subsection{Metrics}
\label{sec:metrics}
% AIM: Change subsubsection titles to use the shorthand names for the metrics as they are referenced throughout the rest of the paper (SNR, w_{0}-w_{a} FOM, \Omega_{c}-\sigma_{8} FOM).
% AIM: This should state that the three metrics are evaluated for clustering, lensing, and 3x2pt, so there are really nine metrics, not two!

@@

Metric 1 would be approximately measurable on real data without performing a full cosmology analysis,
whereas metric 2 would be the final output of such an analysis.  They therefore test performance at what
would be multiple stages in the wider pipeline.
% AIM: Sounds dumb, but readers will appreciate a clear statement that higher is better for these metrics.

\subsubsection{Metric 1: Sigal-to-noise}
% AIM: Suggestion: merge Appendix A with this section because this really doesn't stand by itself

Our first metric is the total signal-to-noise ratio of the power spectrum derived from the assigned
$n_i(z)$ values:
\begin{equation}
    S = \mu^{T} C^{-1} \mu
\label{eq:snr}
\end{equation}
where $\mu$ is a stack of the theoretical predictions for the $C_\ell$ spectra for each tomographic 
bin, in turn, and $C$ a Gaussian estimate of the covariance between them, as described in
\citet{takada_jain} and summarized in \autoref{app:theory}.   We compute this metric for lensing alone, clustering alone, and the full 3x2pt combination.

\subsubsection{Metric 2: Figure-of-merit}
% AIM: The title of this subsubsection should reflect that there are two FOM metrics.

@@

% AIM: We must acknowledge GitHub as a platform that enabled this challenge.
% AIM: Suggestion to specify "subclass" and "superclass" up front for extra-thorough clarity.
The challenge was structured as a python library, in which each classification
method was a class.  The abstract base parent class, {\sc Tomographer}, 
and other challenge machinery, performed initialization,
configuration option handling, data loading, and computing derived quantities
such as colours from magnitudes.

@@

% AIM: Include the code names for these dummy/control methods up front
Two control methods were used to test the challenge machinery and metrics.
% AIM: Explicitly put this into the context of a controlled experiment.
% AIM: Make a new paragraph for each control method.
The first method randomly assigned galaxies to each tomographic bin, resulting
in bins which each had the same $n(z)$, on average (the random noise fluctuations
are very small when using this number of galaxies).  Using this method, we expect
that increasing the number of bins should not increase any metric score, since bins are
maximally correlated.  We find this to be true for all our metrics.

@@

\section{Submitted Methods} \label{sec:methods}
% AIM: This should be a subsection of Challenge Design section above since it's really just a table of the submitted methods directing readers to the appendix for any detail.

% AIM: typo: "Twenty five" --> "Twenty-five"
Twenty five methods were submitted to the challenge, including the control methods described above.
Most used one machine learning method of one type or another to perform the primary classification,
rather than trying to perform a full photometric redshift analysis.

% AIM: These should be a pagewidth table, and we should use macros so the method names display consistently throughout paper.
These methods are described in full in \autoref{app:methods}.  Briefly, they were (with the submitting authors listed in brackets):
% \setlength{\itemsep}{0pt}%
\begin{description*}
	\item[\textsc{Random}] the random assignment control method (organizers)
	\item[\textsc{IBandOnly}] the i-band only control method (organizers)
	\item[\textsc{RandomForest}] a collection of individual decision trees (organizers)
	\item[\textsc{LSTM}] a Long Short-Term Memory neural network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{AutokerasLSTM}] an automaticaly configured LSTM variant\footnote{A technical problem, involving conflicting versions of libraries and incompatible hardware, meant that AutoKerasLSTM could not be run in the final challenge.} (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{LGBM}] a Light Gradient Boosted Machine, a histogram-based tree algorthm (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{TCN}] a Temporal Convolution Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{CNN}] a Convolutional Neural Network (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{Ensemble}] a combination of three other network methods (CRB, BMOF, GT, ESC, EJG)
	\item[\textsc{SimpleSOM}] a Self-Organizing Map (AHW)
	\item[\textsc{PQNLD}] an extension of {\sc SimpleSom} combining with template-based redshift estimation (AHW)
	\item[\textsc{ComplexSOM}] an extension of {\sc SimpleSom} with an additional assignment optimization (DA, AHW, AN, EG, AB)
	\item[\textsc{GPzBinning}] a Gaussian Process redshift estimator and binner (PH)
	\item[\textsc{JaxCNN/JaxResnet}] two related CNN-based bin edge optimizers (AP)
	\item[\textsc{NeuralNetwork1/2}] a dense network optimizing bin assignment for two metrics (AP)
	\item[\textsc{PCACluster}] a Principal Component Analysis of fluxes then optimized clustering (DG)
	\item[\textsc{UTOPIA}] a nearest-neighbour method, optimised in the limit of large representative training sets (AHW)
	\item[\textsc{ZotBin/ZotNet}] two related neural network methods with a preprocesing normalizing flow (DK)
	\item[\textsc{FFNN}] a feed-forward neural network (EN)
	\item[\textsc{FunBins}] a random forest with various nominal bin edge selection (AB)
	\item[\textsc{MLPQNA}] a multi-layer perceptron (SC and MB)
	\item[\textsc{Stacked Generalization}] an combination of gradient tree boosting and logistic regression (JEC)
\end{description*}

\section{Results \& Discussion} \label{sec:results}
% AIM: Convert the first two subsections to the preamble of this section.
% AIM: Preamble should include the tl;dr of "winners."

@@

% AIM: Move this table to the appendix.
\begin{table*}[]
	\begin{tabular}{|l|llll|llll|}
		\hline
		& \multicolumn{4}{c|}{\textbf{riz}}      & \multicolumn{4}{c|}{\textbf{griz}}                                \\ \hline
		\textbf{Method} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} & \textbf{3-bin} & \textbf{5-bin} & \textbf{7-bin} & \textbf{9-bin} \\ \hline
		\input{results/tables/table_FOM_DETF_3x2_dc2.tex}
		\hline
	\end{tabular}
	\caption{Values of the 3x2pt DETF ($w_0,w_a$) figure-of-merit achieved by entrant methods on the 
		CosmoDC2 part of the challenge.}
	\label{tab:cosmodc2}
\end{table*}

The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: $griz$ vs. $riz$,
Buzzard vs. CosmoDC2, and 3, 5, 7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}), the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).

\subsection{Overview}
% AIM: Suggestion: Convert this subsection to a bulleted list to give it some structure harking to the detailed subsections to come. The short versions are long enough to come off as unsupported statements rather than previews of what will be demonstrated in subsequent subsections.

The DETF results for each method are shown in Table \ref{tab:cosmodc2}.  Complete results for all metrics
for the nine-bin scenario only are shown in appendix tables \ref{tab:full_dc2} and \ref{tab:full_buzz}.  
% AIM: Use autoref to appendix.
In all these tables, missing entries are shown for two reasons.
The first, marked with an asterisk, is when a method did not run for the given scenario,
either due to memory overflow or taking an extremely long time.
The second, marked with a dash, is when a method ran, but generated at least one bin with an extremely 
small number count, such that the spectrum could not be calculated.

% AIM: This is the first mention of this as a classification problem. We should mention this up in experimental design along with the assumption of ML/training-based methods, but also change "clasified" [SIC] to "assigned" for overall consistency (and maybe "that entrants clasified" --> "entrants to classify," and maybe reword post-semicolon part as "aside from the impact of reduced final sample size on the metrics, no penalty was applied for discarding galaxies entirely").
% AIM: Also, the discussion of non-assignment should be its own subsection.
The challenge did not require that entrants clasified all galaxies; ambiguous
galaxies could be discarded at the cost of number density.  Several methods
took advantage of this.  The ComplexSom method (see \autoref{sec:csom}) found
it did not improve scores, but as noted in \autoref{sec:pqnld} this may be an artifact
of the high density of the training data, and so could be explored in future challenges.

\begin{figure}[htbp]
	% \centering
	\includegraphics[width=1\linewidth]{results/funbins_nz.pdf}
	\caption{The tomographic $n(z)$ values generated by the winning method in the challenge, {\sc funbins},
		generating nine bins with the CosmoDC2 data.
		The upper panel shows results for the method applied to $riz$ bands, and the middle on $griz$. The
		increased spread and overlap of the bins in the middle panel decreases the overall constraining
		power and signal-to-noise.  The lower panel is for comparison and is not part of the challenge,
		and illutrates results using all the $ugrizy$ bands.}
	\label{fig:funbin_nz}
\end{figure}

The overall winning method in the 3x2pt case in the fiducial $griz$ 3x2pt CosmoDC2 metric was \textsc{funbins},
which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin $ugrizy$ bands.
% AIM: Split paragraph here to distinguish topics of overall challenge winner and conclusions about band information limitation.
% AIM: Put the caveat at the beginning of the following sentence rather than the end. Also, this claim is not well-supported even by Fig. 3. . . 
It is clear from this that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded, at least in our idealised scenario.
In both the CosmoDC2 and Buzzard analyses,
% AIM: analyses --> data sets?
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

There is a clear widening of the $n(z)$ when losing the $g$-band.  This widening leads to a loss in 
constraining power, which is discussed more in Section \ref{sec:gband}.

For comparison, we also show results using all six LSST bands, $ugrizy$; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
is 169.7, barely higher than the score for $griz$, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.

@@

@@

\section{Mistakes} \label{sec:mistakes}
% AIM: This should not be an appendix and should instead be part of the Results & Discussion section. They should not be presented as "mistakes" -- they are part of the experimental design and constitute caveats, which are simply facts, not subjective judgements.

The challenge organizers made a number of mistakes when building and running the challenge.
These do not invalidate our process or results, but we describe some of them here in the hope of
being useful for future challenge designers.
